Naturalness in U(1) exentions of the MSSM

### 2014-11-10 paper/e6ssm-tuning-paper.tex
added more references and some E6SSM literature review.

U(1) extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{} and also raise the Higgs mass with new $F$- and
$D$-terms. Nonetheless, as was recently demonstrated in the context of
the exceptional supersymmetric standard model (E$_6$SSM)
\cite{King:2005jy,King:2005my,Athron:2010zz}, can still suffer from
naturalness problems with the mass of the new $Z^\prime$ associated
with break down of the new U(1) appearing in the electroweak symmetry
breaking conditions at tree level \cite{Athron:2013ipa}.  Nonetheless
the constrained version of E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue,
  Athron:2009bs} still turned out to be significantly less tuned than
the cMSSM.

@@

U(1) literature review  \newline

E6 literature review \newline
 Exceptional Supersymmetric Standard Model is a very well U(1) extension where....

>>>>

U(1) extensions of the MSSM provide a very elegant solution to this
$\mu$-problem \cite{Fayet:1977yc, Kim:1983dt, Suematsu:1994qm, Cvetic:1995rj,  Cvetic:1996mf, Jain:1995cb, Nir:1995bu, Cvetic:1997ky} and also
raise the Higgs mass with new $F$- and $D$-terms. Nonetheless, as was
recently demonstrated in the context of the exceptional supersymmetric
standard model (E$_6$SSM)
\cite{King:2005jy,King:2005my,Athron:2010zz}, can still suffer from
naturalness problems with the mass of the new $Z^\prime$ associated
with break down of the new U(1) appearing in the electroweak symmetry
breaking conditions at tree level \cite{Athron:2013ipa}.  Nonetheless
the constrained version of E$_6$SSM (cE$_6$SSM) \cite{Athron:2009ue,
  Athron:2009bs} still turned out to be significantly less tuned than
the cMSSM.

@@

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem as the USSM
\cite{Cvetic:1995rj, Jain:1995cb, Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}, while we will
sometimes specify that the models are inspired by $E_6$ by referring
to them as such. Finally we will also discuss specific models with
complete $E_6$ matter multiplets at low energies and an extra $U(1)$
under which the right handed neutrino remains interactionless as
E$_6$SSM-variants\cite{Howl:2007zi, Braam:2009fi,
  Braam:2010sy, Hall:2011zq, Nevzorov:2012hs, Athron:2014pua}, with the E$_6$SSM reserved for the original
formualtion of this model \cite{King:2005jy,King:2005my,Athron:2010zz}
original formulation of the model.

$U(1)$ and $E_6$ inspired extensions of the MSSM have been studied
very widely in the literature \cite{Gunion:1989we, Gunion:1992hs,
  Binetruy:1985xm, Ellis:1986yg, Ibanez:1986si, Gunion:1986ky,
  Haber:1986gz, Baer:1987eb, Gunion:1987jd, Grifols:1986vr,
  Ellis:1986ip, Morris:1987fm, Drees:1987tp, Ma:1995xk,
  Suematsu:1997tv, Suematsu:1997qt, Suematsu:1997au, Keith:1996fv,
  Keith:1997zb, Gherghetta:1996yr, Demir:1998dk,L angacker:1998tc,
  Hambye:2000bn, Ma:2000jf} (or for reviews see
\cite{Hewett:1988xc,Langacker:2008yv}).  There has also been a lot of
work recently including investigations of the neutralino sector
\cite{Hesselbach:2001ri, Barger:2005hb, Choi:2006fz,
  Barger:2007nv};the relic abundance of dark matter
\cite{Kalinowski:2008iq}; GUT scale family symmetries which can
explain the hierarchy of masses in the fermion sector and their
associated mixings \cite{Stech:2008wd}; neurtrino physics
\cite{Kang:2004ix}; explanations of the matter-anti-matter asymmetry
of the universe though EW baryogensis or leptogensis
\cite{Hambye:2000bn,Ma:2000jf,Kang:2004pp}; dipole moments
\cite{GutierrezRodriguez:2006hb}; anomaly mediated SUSY breaking with
D-term contributions \cite{Asano:2008ju} and the (extended) Higgs
sectors \cite{Daikoku:2000ep,Ham:2008xf}.

In this paper we will focus a lot on a very well motivated $E_6$
inspired $U(1)$ exension, the Exceptional Supersymmetric Standard
Model (E$_6$SSM)\cite{King:2005jy,King:2005my,Athron:zz}.  In this
model since the right handed neutrino remains interactionless they may
explain the tiny observed masses of neutrinos via the see-saw
mechanism andp the baryon asymmetry in the universe via leptogensis
\cite{Hambye:2000bn,King:2008qb}.

The gauge coupling running in the E$_6$SSM at the two-loop level leads
to unification more precisely than in the MSSM \cite{King:2007uj} or
in slightly modified scenarios two-step unification can take place
\cite{Howl:2007hq,Howl:2007zi}. If the exotic particles are light in
these models this can open up non-standard decays of the SM--like
Higgs boson \cite{Hall:2010ix,Nevzorov:2013tta,Athron:2014pua}. The
correct relic abundance could be obtained entirely through an almost
decoupled ``inert'' neutralino sector \cite{Hall:2009aj} no longer
seem phenomenologically viable due to limits from direct detection of
dark matter \cite{} and limits on invisible Higgs decays \cite{}.

Alternatively one may specilaize to scenarios known as the EZSSM
\cite{Hall:2011zq} where the inert singlinos, which cause these
problems are entirely decoupled and the relic abundance is fitted with
a bino-like candidate with a novel mechanism involving back-scattering
into a Heavier inert Higgsino. Finally one may obtain...

In the
constrained version of the of the E$_6$SSM the particle spectrum,
collider signatures and fine tuning have been studied
\cite{cE6SSM1,cE6SSM2,cE6SSM3,cE6SSM4,Athron:2013ipa}. The threshold
corrections to the $\overline{DR}$ gauge and Yukawa couplings in the
\essm were calculated and the numerical impact in the constrained
version examined \cite{Athron:2012pw}. The impact of gauge kinetic
mixing in the case where both of the extra U(1)'s appearing from the
breakdown of E$_6$ are present at low energy was studied in
\cite{Rizzo:2012rf}.  The \essm was also included in a studies looking
at how first or second generation sfermion masses can be used to
constrain the GUT scale parameters \cite{Miller:2012vn} and the
renormalization of VEVs \cite{Sperling:2013-1, Sperling:2013-2}.
